---
title: "FEWSNET_processing"
author: "Ferdinand Seyffer"
date: "2025-10-07"
output:
  html_document:
    code_folding: hide
    toc: true
    number_sections: true
    toc_float:
      collapsed: true
    theme: cerulean
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Load necessary packages, if not installed install with "install.packages()
```{r message=F, warning=F}
library(sf)
library(dplyr)
library(stringr)
library(ggplot2)
library(purrr)
library(tidyr)
library(scales)
```

Setup data
```{r message=F, warning=F}
  # Turn off s2 fo planar instead of spherical geodata handling
  sf_use_s2(FALSE)  
  
  # Load Somalia districts
  districts <- st_read("Admin/som_adm_ocha_20250108_ab_shp/som_admbnda_adm2_ocha_20250108.shp")
  
  # List and load FEWSNET shapefiles
  foodsec_files <- list.files(path = "FI_data/FEWSNET_EA_all/", pattern = "\\.shp$",  full.names = TRUE)
```


## Preprocessing

Loop through FEWSNET files, intersect with ADM2
```{r message=F, warning=F}
 intersections <- list()
  
  for (fp in foodsec_files) {
    food_sec <- st_read(fp, quiet = TRUE) # load food security file
    
    # Reproject to CRS of districts file
    if (is.na(st_crs(food_sec))) st_crs(food_sec) <- 4326
    food_sec <- st_transform(food_sec, 4326)
    
    # Fix geometries because of common invalid geometries with FEWSNET data
    food_sec <- st_make_valid(food_sec)
    
    # Intersect FS file with districts, keeping only column "CS" from FS layer
    intersection <- st_intersection(districts, food_sec["CS"])
    
    # extract base name (without .shp) for both naming + date
    basename <- tools::file_path_sans_ext(basename(fp))
    
    # add date column based on basename of FS files
    intersection <- intersection %>%
      mutate(
        source = basename,
        date = str_extract(basename, "\\d{6}") %>%
          paste0("01") %>%
          as.Date("%Y%m%d")
      )
    
    # Add each file to intersections list
    intersections[[basename]] <- intersection
  }
```


Convert distinct intersections to one dataframe
```{r message=F, warning=F}
  all_intersections <- map_dfr(intersections, \(x) st_drop_geometry(x))|> # drop geometries and combine to dataframe
    distinct(ADM2_EN, CS, date, .keep_all = TRUE) |> # keep just rows where ADM2_EN, CS and date are unique
   
    # Add TRUE/FALSE columns for different conditions
     mutate(
      cs3 = CS == 3,    # CS = 3
      cs4 = CS == 4,    # CS = 4
      cs5 = CS == 5,    # CS = 5
      cs_gr2 = CS >= 3, # CS over 2
      cs_gr3 = CS >= 4  # CS over 3
    )
  
```


Convert distinct intersections to one dataframe
```{r message=F, warning=F}
  clean_intersections <- all_intersections |>
    # drop invalid/missing values
    filter(CS != 99) |>
    # keep only selected columns
    select(ADM2_EN, date, ADM2_PCODE, CS, cs3, cs4, cs5, cs_gr2, cs_gr3)

```


Group by districts and calculate different basic stats
```{r message=F, warning=F}
  fewsnet_stats <- clean_intersections |>
    group_by(ADM2_PCODE, ADM2_EN) |>       # group by districts
    
    # Summarise per districts: times in specific cs, times in cs >2/ >3
    summarise(
      sum_cs3   = sum(cs3,   na.rm = TRUE),
      sum_cs4   = sum(cs4,   na.rm = TRUE),
      sum_cs5   = sum(cs5,   na.rm = TRUE),
      sum_cs_gr2 = sum(cs_gr2, na.rm = TRUE),
      sum_cs_gr3 = sum(cs_gr3, na.rm = TRUE),
      mean_CS   = mean(CS,   na.rm = TRUE),
      .groups = "drop"
    )
```

Join ipc stats back to district geometries (e.g. for import in QGIS)
```{r message=F, warning=F}
  # Join data back to geometries
  fewsnet_stats_sf <- districts |>
    select(ADM2_PCODE, geometry) |>   # keep only join key + geometry
    left_join(fewsnet_stats, by = "ADM2_PCODE") |>
    st_as_sf()
  
  # Write file 
  st_write(fewsnet_stats_sf,
           "FEWSNET_adm2_stats.gpkg",
           delete_dsn = TRUE) 
```


##  Data Analysis

Timeseries plots for IPC for four selected districts
```{r message=F, warning=F}
  # Enter districts of interest here
  sel_districts <- c("Luuq", "Xudur", "Xarardheere", "Hobyo")
  
  # Summarise per district & month
  district_summary <- clean_intersections %>%
    filter(ADM2_EN %in% sel_districts) %>%
    group_by(ADM2_EN, date) %>%
    summarise(
      mean_CS = mean(CS, na.rm = TRUE),
      max_CS  = max(CS, na.rm = TRUE),
      .groups = "drop"
    )
  
  # Calculate overall CS mean per district (for dashed mean line)
  overall_means <- district_summary %>%
    group_by(ADM2_EN) %>%
    summarise(overall_mean = mean(mean_CS, na.rm = TRUE))
  
  # Plot
  ggplot(district_summary, aes(x = date, group = ADM2_EN)) +
    geom_line(aes(y = mean_CS, color = "Mean CS"), linewidth = 1) +
    geom_line(aes(y = max_CS,  color = "Max CS"),  linewidth = 1, linetype = "dashed") +
    geom_hline(data = overall_means, aes(yintercept = overall_mean),
               linetype = "dashed", color = "black") +
    facet_wrap(~ ADM2_EN, ncol = 2) +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    scale_color_manual(values = c("Mean CS" = "steelblue", "Max CS" = "red3")) +
    labs(x = "Date", y = "CS", color = "Statistic") +
    theme_minimal()
```


Barcharts for times in IPC ≥3/4 & mean IPC for all districts
```{r message=F, warning=F}
 # --- CS ≥ 3 ---
  ggplot(fewsnet_stats, aes(x = reorder(ADM2_EN, sum_cs_gr2))) +
    geom_col(aes(y = sum_cs_gr2), fill = "red3") +
    geom_line(aes(y = rescale(mean_CS, to = c(0, max(sum_cs_gr2)), from = c(2, 3)), group = 1),
              color = "black") +
    geom_point(aes(y = rescale(mean_CS, to = c(0, max(sum_cs_gr2)), from = c(2, 3))),
               color = "black") +
    scale_y_continuous(
      name = "Count of CS ≥ 3",
      sec.axis = sec_axis(~rescale(., from = c(0, max(fewsnet_stats$sum_cs_gr2)), to = c(2, 3)),
                          name = "Mean CS")
    ) +
    coord_flip() +
    labs(title = "Districts with CS ≥ 3") +
    theme_minimal()
  
  # --- CS ≥ 4 ---
  ggplot(fewsnet_stats, aes(x = reorder(ADM2_EN, sum_cs_gr3))) +
    geom_col(aes(y = sum_cs_gr3), fill = "steelblue") +
    geom_line(aes(y = rescale(mean_CS, to = c(0, max(sum_cs_gr3)), from = c(2, 3)), group = 1),
              color = "black") +
    geom_point(aes(y = rescale(mean_CS, to = c(0, max(sum_cs_gr3)), from = c(2, 3))),
               color = "black") +
    scale_y_continuous(
      name = "Count of CS ≥ 4",
      sec.axis = sec_axis(~rescale(., from = c(0, max(fewsnet_stats$sum_cs_gr3)), to = c(2, 3)),
                          name = "Mean CS")
    ) +
    coord_flip() +
    labs(title = "Districts with CS ≥ 4") +
    theme_minimal()
```

Histogramms for districts in IPC ≥3/4 over whole time period
```{r message=F, warning=F}
  # group by date to get district nr. per period
  counts <- clean_intersections %>%
    group_by(date) %>%
    summarise(
      n_cs3plus = sum(cs_gr2, na.rm = TRUE),
      n_cs4plus = sum(cs_gr3, na.rm = TRUE),
      .groups = "drop"
    )
  
  # plot for CS ≥3
  ggplot(counts, aes(x = date, y = n_cs3plus)) +
    geom_col(fill = "red3") +
    labs(title = "Districts in CS ≥ 3", x = "Date", y = "Count") +
    theme_minimal()
  
  # plot for CS ≥4
  ggplot(counts, aes(x = date, y = n_cs4plus)) +
    geom_col(fill = "steelblue") +
    labs(title = "Districts in CS ≥ 4", x = "Date", y = "Count") +
    theme_minimal()
```



