---
title: "IPC_processing"
author: "Ferdinand Seyffer"
date: "2025-10-06"
output:
  html_document:
    code_folding: hide
    toc: true
    number_sections: true
    toc_float:
      collapsed: true
    theme: cerulean
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Load necessary packages, if not installed install with "install.packages()
```{r message=F, warning=F}
library(sf)
library(dplyr)
library(stringr)
library(ggplot2)
library(purrr)
library(tidyr)
library(scales)
```

Setup data
```{r message=F, warning=F}
# Turn off s2 fo planar instead of spherical geodata handling
sf_use_s2(FALSE)  


# Load Somalia districts
districts <- st_read("Admin/som_adm_ocha_20250108_ab_shp/som_admbnda_adm2_ocha_20250108.shp")

# List and load FEWSNET shapefiles
foodsec_files <- list.files(path = "FI_data/IPC_all/", pattern = "\\.geojson$",  full.names = TRUE)
```


## Preprocessing

Loop through FEWSNET files, intersect with ADM2
```{r message=F, warning=F}
intersections <- list()

for (fp in foodsec_files) {
  food_sec <- st_read(fp, quiet = TRUE) # load food security file
  
  # Reproject to CRS of districts file
  if (is.na(st_crs(food_sec))) st_crs(food_sec) <- 4326
  food_sec <- st_transform(food_sec, 4326)
  
  # Fix geometries because of common invalid geometries with FEWSNET data
  food_sec <- st_make_valid(food_sec)
  
  # Intersect FS file with districts, keeping only column "overall_phase" from FS layer
  intersection <- st_intersection(districts, food_sec["overall_phase"])
  
  # extract base name (without .shp) for both naming + date
  basename <- tools::file_path_sans_ext(basename(fp))
  
  # add date column based on basename of FS files
  intersection <- intersection %>%
    mutate(
      source = basename,
      # extract "August_2022" (month name + underscore + year)
      month_year = str_extract(basename, "[A-Za-z]+_[0-9]{4}"),
      # replace underscore with space -> "August 2022"
      month_year = gsub("_", " ", month_year),
      # parse into a Date (assume day = 01)
      date = as.Date(paste0("01 ", month_year), format = "%d %B %Y")
    )
  
  # Add each file to intersections list
  intersections[[basename]] <- intersection
}
```


Convert distinct intersections to one dataframe
```{r message=F, warning=F}
all_intersections <- map_dfr(intersections, \(x) st_drop_geometry(x))|># drop geometries and combine to dataframe
  distinct(.data[["ADM2_EN"]], ADM2_EN, overall_phase, date, .keep_all = TRUE) |> # keep just rows where ADM2_EN, overall_phase and date are unique
  
  # Add TRUE/FALSE columns for different conditions
  mutate(
    ipc3 = overall_phase == 3,    # overall_phase = 3
    ipc4 = overall_phase == 4,    # overall_phase = 4
    ipc5 = overall_phase == 5,    # overall_phase = 5
    ipc_gr2 = overall_phase >= 3, # overall_phase over 2
    ipc_gr3 = overall_phase >= 4  # overall_phase over 3
  )
```


Convert distinct intersections to one dataframe
```{r message=F, warning=F}
clean_intersections <- all_intersections |>
  # drop invalid/missing values
  filter(overall_phase != 99) |>
  # keep only selected columns
  select(ADM2_EN, date, ADM2_PCODE, overall_phase, ipc3, ipc4, ipc5, ipc_gr2, ipc_gr3)

```


Group by districts and calculate different basic stats
```{r message=F, warning=F}
ipc_stats <- clean_intersections |>
  group_by(ADM2_PCODE, ADM2_EN) |>       # group by districts
  
  # Summarise per districts: times in specific IPC, times in IPC >2/ >3
  summarise(
    sum_ipc3   = sum(ipc3,   na.rm = TRUE),
    sum_ipc4   = sum(ipc4,   na.rm = TRUE),
    sum_ipc5   = sum(ipc5,   na.rm = TRUE),
    sum_ipc_gr2 = sum(ipc_gr2, na.rm = TRUE),
    sum_ipc_gr3 = sum(ipc_gr3, na.rm = TRUE),
    mean_ipc   = mean(overall_phase,   na.rm = TRUE),
    .groups = "drop"
  )

```

Join ipc stats back to district geometries (e.g. for import in QGIS)
```{r message=F, warning=F}
# Join data back to geometries
ipc_stats_sf <- districts |>
  select(ADM2_PCODE, geometry) |>   # keep only join key + geometry
  left_join(ipc_stats, by = "ADM2_PCODE") |>
  st_as_sf()

# Write file 
st_write(ipc_stats_sf,
         "ipc_adm2_stats.gpkg",
         delete_dsn = TRUE) 

# Keep a copy of IPC results
ipc_clean <- clean_intersections
ipc_stats_final <- ipc_stats
```


##  Data Analysis

Timeseries plots for IPC for four selected districts
```{r message=F, warning=F}
# Enter districts of interest here
sel_districts <- c("Luuq", "Xudur", "Xarardheere", "Hobyo")

# Summarise per district & month
district_summary <- clean_intersections %>%
  filter(ADM2_EN %in% sel_districts) %>%
  group_by(ADM2_EN, date) %>%
  summarise(
    mean_ipc = mean(overall_phase, na.rm = TRUE),
    max_IPC  = max(overall_phase, na.rm = TRUE),
    .groups = "drop"
  )

# Calculate overall IPC mean per district (for dashed mean line)
overall_means <- district_summary %>%
  group_by(ADM2_EN) %>%
  summarise(overall_mean = mean(mean_ipc, na.rm = TRUE))

# Plot
ggplot(district_summary, aes(x = date, group = ADM2_EN)) +
  geom_line(aes(y = mean_ipc, color = "Mean IPC"), linewidth = 1) +
  geom_line(aes(y = max_IPC,  color = "Max IPC"),  linewidth = 1, linetype = "dashed") +
  geom_hline(data = overall_means, aes(yintercept = overall_mean),
             linetype = "dashed", color = "black") +
  facet_wrap(~ ADM2_EN, ncol = 2) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_color_manual(values = c("Mean IPC" = "steelblue", "Max IPC" = "red3")) +
  labs(x = "Date", y = "IPC", color = "Statistic") +
  theme_minimal()

```


Barcharts for times in IPC ≥3/4 & mean IPC for all districts
```{r message=F, warning=F}
# --- IPC ≥ 3 ---
ggplot(ipc_stats, aes(x = reorder(ADM2_EN, sum_ipc_gr2))) +
  geom_col(aes(y = sum_ipc_gr2), fill = "red3") +
  geom_line(aes(y = rescale(mean_ipc, to = c(0, max(sum_ipc_gr2)), from = c(2, 3)), group = 1),
            color = "black") +
  geom_point(aes(y = rescale(mean_ipc, to = c(0, max(sum_ipc_gr2)), from = c(2, 3))),
             color = "black") +
  scale_y_continuous(
    name = "Count of IPC ≥ 3",
    sec.axis = sec_axis(~rescale(., from = c(0, max(ipc_stats$sum_ipc_gr2)), to = c(2, 3)),
                        name = "Mean IPC")
  ) +
  coord_flip() +
  labs(title = "Districts with IPC ≥ 3") +
  theme_minimal()

# --- IPC ≥ 4 ---
ggplot(ipc_stats, aes(x = reorder(ADM2_EN, sum_ipc_gr3))) +
  geom_col(aes(y = sum_ipc_gr3), fill = "steelblue") +
  geom_line(aes(y = rescale(mean_ipc, to = c(0, max(sum_ipc_gr3)), from = c(2, 3)), group = 1),
            color = "black") +
  geom_point(aes(y = rescale(mean_ipc, to = c(0, max(sum_ipc_gr3)), from = c(2, 3))),
             color = "black") +
  scale_y_continuous(
    name = "Count of IPC ≥ 4",
    sec.axis = sec_axis(~rescale(., from = c(0, max(ipc_stats$sum_ipc_gr3)), to = c(2, 3)),
                        name = "Mean IPC")
  ) +
  coord_flip() +
  labs(title = "Districts with IPC ≥ 4") +
  theme_minimal()
```

Histogramms for districts in IPC ≥3/4 over whole time period
```{r message=F, warning=F}
# group by date to get district nr. per period
counts <- clean_intersections %>%
  group_by(date) %>%
  summarise(
    n_ipc3plus = sum(ipc_gr2, na.rm = TRUE),
    n_ipc4plus = sum(ipc_gr3, na.rm = TRUE),
    .groups = "drop"
  )

# plot for IPC ≥3
blabla = ggplot(counts, aes(x = date, y = n_ipc3plus)) +
  geom_col(fill = "red3") +
  labs(title = "Districts in IPC ≥ 3", x = "Date", y = "Count") +
  theme_minimal()

# plot for IPC ≥4
ggplot(counts, aes(x = date, y = n_ipc4plus)) +
  geom_col(fill = "steelblue") +
  labs(title = "Districts in IPC ≥ 4", x = "Date", y = "Count") +
  theme_minimal()


```



