---
title: "Sketch Map Tool Numbers"
author: "Team Sketch Map Tool"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    number_sections: true
    toc_float:
      collapsed: true
    theme: cerulean
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(ggplot2)
library(lubridate)
library(plotly)
library(tmap)
library(sf)
library(rnaturalearth)
library(ggVennDiagram)
library(ggmosaic)


```
# Background and data

Since 2024-12-05 we are analyzing the Apache Log which provides us with insides of the tool usage based in the IP adress.


# Creation of Map Frames



```{r, echo=FALSE, message=FALSE, warning=FALSE}

#load the data. Ne newest csv files will be saved in a folder, which should always only contain the three files.

files <- list.files("SMT_Input/", full.names = TRUE)

if (length(files) >= 3) {
  first_file <- files[1]
  second_file <- files[2]
  third_file <- files[3]
  
  # Read files using read_delim (or read.csv if needed)
  sketchmap_report_map_frames <- read_delim(first_file, delim = ";", escape_double = FALSE, trim_ws = TRUE)
  sketchmap_report_sketches <- read_delim(second_file, delim = ";", escape_double = FALSE, trim_ws = TRUE)
  sketchmap_report_uploads <- read_delim(third_file, delim = ";", escape_double = FALSE, trim_ws = TRUE)
} else {
  stop("Not enough files in the folder.")
}

#sketchmap_report_uploads_20250215 <- read_delim("SMT_Input/sketchmap_report_uploads_20250215.csv", 
   # delim = ";", escape_double = FALSE, trim_ws = TRUE)
#View(sketchmap_report_uploads_20250215)


#sketchmap_report_sketches_20250215 <- read_delim("SMT_Input/sketchmap_report_sketches_20250215.csv", 
   # delim = ";", escape_double = FALSE, trim_ws = TRUE)
#View(sketchmap_report_sketches_20250215)

#sketchmap_report_map_frames_20250215 <- read_delim("SMT_Input/sketchmap_report_map-frames_20250215.csv", 
   # delim = ";", escape_double = FALSE, trim_ws = TRUE)
#View(sketchmap_report_map_frames_20250215)

sketchmap_report_map_frames_filter <- sketchmap_report_map_frames |> filter(created_at > '2024-12-05' |> as.Date())
# cum counts per day for each layer
cumulative_counts <- sketchmap_report_map_frames_filter |>
  arrange(created_at) |>
  mutate(day = floor_date(created_at, "day")) |>
  count(day, layer) |>
  group_by(layer) |>
  arrange(day) |>  # Ensure correct order within each layer
  mutate(cumulative_count = cumsum(n)) |>
  ungroup()

# cum counts per day for all sketch maps
total_counts <- cumulative_counts |>
  group_by(day) |>
  summarise(n = sum(n), .groups = "drop") |>
  mutate(layer = "Total") |>
  arrange(day) |>
  mutate(cumulative_count = cumsum(n))

# combine to one long table
final_counts <- bind_rows(cumulative_counts, total_counts) |>
  arrange(day, layer)


# plot
chart1 <- final_counts |>
  ggplot(aes(x = day, y = cumulative_count, color=layer)) +
  geom_line() +
  geom_point() +
  labs(title = "Cumulative Count of Mapframes Over Time",
       x = "Days",
       y = "Cumulative Count") +
  theme_classic()

ggplotly(chart1)
```

# Bounding Boxes
```{r,message=FALSE}
#switch to interactive
tmap_mode("view")

mapframes_bboxes <- st_as_sf(sketchmap_report_map_frames, wkt = "bbox", crs = 4326)

tm_shape(mapframes_bboxes) +
  tm_basemap("OpenStreetMap") +
  tm_borders() +
  tm_layout(title = "Bounding Box Localization for Generated Map Frames")


```



# Bounding Boxes per Country

```{r,message=FALSE}

#areas of bounding boxes: not sure yet if of interest
#mapframes_bboxes_proj <- st_transform(mapframes_bboxes, crs = 32633)

#calculate areas
#mapframes_bboxes_proj$area <- st_area(mapframes_bboxes_proj)


# Calculate centroids of the bounding boxes

mapframes_bboxes_sf <- st_as_sf(mapframes_bboxes, wkt = "bbox", crs = 4326)
mapframes_bboxes_sf$centroid <- st_centroid(mapframes_bboxes_sf)

#Spatial join with world
world <- ne_countries(scale = "medium", returnclass = "sf")

centroids_with_country <- st_join(mapframes_bboxes_sf$centroid , world)

#Groupby
centroid_counts <- centroids_with_country %>%
  group_by(iso_a2) %>%
  summarise(countbbox = n_distinct(centroid), .groups = "drop")
#join back
centroid_counts_sf <- st_join(world, centroid_counts)

tmap_mode("view")

tm_shape(centroid_counts_sf ) + 
  tm_borders() +  # Add borders of countries
  tm_fill(col = "countbbox", 
          palette = "brewer.reds", 
          title = "Centroid Count",
          style = "jenks",  # Change classification method 
          na.color = "gray",
          n = 5) +  # Number of classes for Jenks classification
  tm_layout(title = "Map Frames created for Countries")

```



# Downloaded Sketch Maps

```{r,message=FALSE}

#How many created SKetch Maps were actually dowloaded?
sketchmap_report_map_frames_filter |> filter(!is.na(downloaded_at))|> nrow()


```
#Mapframes with downloads and uploads (can be multiple uploads per mapframe)

```{r,message=FALSE}

#any possibility to track multiple uploads?

# Number of downloaded Sketch Maps vs. created Sketch Maps
sketchmap_report_map_frames_filter |>  filter(!is.na(downloaded_at) & uploads>0) |> nrow()
#67
#67 downloads mapframes show and uploads. Uploads csv = 296. Does that mean that there were several multiple uploads???
#im Schnitt 4,5 auf einmal upgeloaded?

#general number of uploads
no_uploads <- nrow(sketchmap_report_uploads)
#296

```

#How was often consent was given to use uploaded sketchmaps to contribute to model training?


```{r,message=FALSE}

sketchmap_report_map_frames_filter |>  filter(!is.na(downloaded_at) & uploads>0 & consenses>0) |> nrow()
#24

test <- sketchmap_report_map_frames_filter |>  filter(!is.na(downloaded_at) & uploads>0)
test$uploads |> sum()
round(test$consenses |> sum() / test$uploads |> sum() *100,2)


```


# Uploaded Sketch Maps and share of dowloaded results

```{r,message=FALSE}

sketchmap_report_uploads |> filter(is.na(download_raster) & is.na(download_vector)) |> nrow()
#22

download_raster <- sketchmap_report_uploads |> filter(!is.na(download_raster))
no_raster_downloads <- nrow(download_raster) 
# 141
share_raster_downloads <- no_raster_downloads / no_uploads

download_raster$uuid |> unique() |> length()
#  79
download_vector <- sketchmap_report_uploads |> filter(!is.na(download_vector))
no_vector_downloads <- nrow(download_vector) 
# 230

share_vector_downloads <- no_vector_downloads / no_uploads

download_vector$uuid |> unique() |> length()
# 138

#interesting that it is from different uuids?

#try a Venn diagram

#sketchmap_report_uploads_20250115 <- sketchmap_report_uploads_20250115 |> 
  #mutate(has_raster = !is.na(download_raster),
         #has_vector = !is.na(download_vector))
#enn_counts <- with(sketchmap_report_uploads_20250115, table(has_raster, has_vector))

#venn_data <- list(
  #Raster = which(sketchmap_report_uploads_20250115$has_raster),
  #Vector = which(sketchmap_report_uploads_20250115$has_vector)
#)
#ggVennDiagram(venn_data)

# Bar Chart 


# Categorize downloads based on raster and vector presence
sketchmap_report_uploads <- sketchmap_report_uploads |> 
  mutate(category = case_when(
    !is.na(download_raster) & !is.na(download_vector) ~ "Both",
    !is.na(download_raster) ~ "Raster Only",
    !is.na(download_vector) ~ "Vector Only",
    TRUE ~ "None"  # This covers cases where neither raster nor vector is downloaded
  ))

# Count each category
category_counts <- table(sketchmap_report_uploads$category)
total <- sum(category_counts) 

# Convert to data frame for ggplot
category_df <- as.data.frame(category_counts)
colnames(category_df) <- c("Category", "Count")

# Calculate percentage
category_df$Percentage <- (category_df$Count / total) * 100

# Plot bar chart with percentages
ggplot(category_df, aes(x = Category, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity") +
  labs(title = "Distribution of Raster and Vector Geodata Downloads", 
       x = "Download Type", 
       y = "Percentage") +
  theme_minimal()



```

# SHare of Opt-in used


# Use of baselayers

```{r,message=FALSE}
#osm esri chart

donut <-sketchmap_report_map_frames_filter |> 
  group_by(layer) |> 
  summarise(count=n())

donut$fraction <- donut$count / sum(donut$count)
donut$ymax <- cumsum(donut$fraction)
donut$ymin <- c(0, head(donut$ymax, n=-1))


chart2 <- ggplot(donut, aes(ymax=ymax, ymin=ymin, xmax=2, xmin=1, fill=layer)) +
  geom_rect() +
  coord_polar(theta="y") +
  xlim(c(0.2, 2.5)) + 
  theme_void() +  
  theme(legend.position = "right") +
  labs(fill="Layer")

chart2

```

# Map format used for map creation

```{r,message=FALSE}
donut_format <-sketchmap_report_map_frames_filter |> 
  group_by(format) |> 
  summarise(count=n())

donut_format$fraction <- donut_format$count / sum(donut_format$count)
donut_format$ymax <- cumsum(donut_format$fraction)
donut_format$ymin <- c(0, head(donut_format$ymax, n=-1))


chart_format <- ggplot(donut_format, aes(ymax=ymax, ymin=ymin, xmax=2, xmin=1, fill=format)) +
  geom_rect() +
  coord_polar(theta="y") +
  xlim(c(0.2, 2.5)) + 
  theme_void() +  
  theme(legend.position = "right") +
  labs(fill="Layer")

chart_format


```


# Orientation chosen for map creation

```{r,message=FALSE}

#why is this not binary and other languages included

donut_orientation <-sketchmap_report_map_frames_filter |> 
  group_by(orientation) |> 
  summarise(count=n())

donut_orientation$fraction <- donut_orientation$count / sum(donut_format$count)
donut_orientation$ymax <- cumsum(donut_orientation$fraction)
donut_orientation$ymin <- c(0, head(donut_orientation$ymax, n=-1))


chart_orientation <- ggplot(donut_orientation, aes(ymax=ymax, ymin=ymin, xmax=2, xmin=1, fill=orientation)) +
  geom_rect() +
  coord_polar(theta="y") +
  xlim(c(0.2, 2.5)) + 
  theme_void() +  
  theme(legend.position = "right") +
  labs(fill="Layer")

chart_orientation
```



# Language used to create Sketch Map

```{r, message = FALSE}

#why are there ES and es, why are there missing values if we have filtered already to dec 5

donut_lang <-sketchmap_report_map_frames_filter |> 
  group_by(created_lang) |> 
  summarise(count=n())

donut_lang$fraction <- donut_lang$count / sum(donut_lang$count)
donut_lang$ymax <- cumsum(donut_lang$fraction)
donut_lang$ymin <- c(0, head(donut_lang$ymax, n=-1))

# Total count of all values(to put to the legend, does not work so far)
total_count <- sum(donut_lang$count)

chart_lang <- ggplot(donut_lang, aes(ymax=ymax, ymin=ymin, xmax=2, xmin=1, fill=created_lang)) +
  geom_rect() +
  coord_polar(theta="y") +
  xlim(c(0.2, 2.5)) + 
  theme_void() +  
  theme(legend.position = "right") +
  labs(fill="Layer") +
  annotate("text", x = 2.7, y = 0, label = paste("Total: ", total_count), size = 5,hjust = 0)

chart_lang
```


# Countries where Sketch Maps have been created from

```{r,message=FALSE}

plot_data_country <- sketchmap_report_map_frames_filter %>%
  group_by(created_from_country) %>%
  summarise(n = n())

#why are there entries since dec 5 without information created_from_country (as well as no information for other columns based on Apache)


#download worldmap
world <- ne_countries(scale = "medium", returnclass = "sf")
#world_lokal <- st_read("/Users/aschauss/Documents/SketchMapTool/Statistics/Map/world-administrative-boundaries.geojson")


#transfrom to Mollweide
world_mollweide <- st_transform(world, crs = "+proj=moll")
#world_lokal <- st_transform(world_lokal, crs = "+proj=moll")

#group by iso code
sketchmap_report_map_frames_filter_iso <- sketchmap_report_map_frames_filter %>%
  group_by(created_from_iso_code) %>%
  summarise(count = sum(!is.na(created_from_iso_code)))

#rightjoin with world
sketchmap_report_world_iso <- sketchmap_report_map_frames_filter_iso %>%
  rename(iso_a2 = created_from_iso_code) %>%
  right_join(world_mollweide, by = "iso_a2")

#back to sf
sketchmap_report_world_iso <- st_as_sf(sketchmap_report_world_iso)

tmap_mode("view")

#print map vr
tm_shape(sketchmap_report_world_iso) +
  tm_polygons("count", 
              palette = "brewer.reds",
              title = "Count per Country",
              style = "jenks",  # Change classification method
              n = 5  # Number of classes
              ) +
  tm_borders() +
  tm_layout(main.title = "Created Sketch Maps",
            legend.outside = TRUE,
            legend.position = c("left", "bottom"),
            legend.bg.color = "white",
            legend.frame = FALSE,
            frame = FALSE
  )


```


